Zohmg---a large scale data store for aggregated time-series-based data. [DRAFT]

### 2009-08-15 doc/report/msc-report.tex
6.2 - discussion

Discussion here.

>>>>

The decision to implement the import functionality on Hadoop was driven by both
practical and environmental reasons: it is unfeasible to analyse large data sets
on a non-distributed system so some sort of distributed computation model was
needed to reach a satisfiable performance, and Last.fm already has a large
Hadoop cluster and a technology culture in which Hadoop is used extensively. The
decision to use Hadoop was in the end a non-decision, an already made
conclusion.

The reasons behind using Dumbo and writing Zohmg in Python are similar. Last.fm
was already moving towards writing a larger set of its analysis tools in Python
thanks to Dumbo, which is a product of another intern at the company. Python is
a language which allows for rapid prototyping and is as such well suited for
writing both the user's map function and indeed the whole of Zohmg.

The main point of contention was which backing store to use. A number of
different systems that fitted the requirements were tried out but none of them
proved to be either very dissimilar from the BigTable apporach of HBase, or
indeed ready for production. HBase has the advantage of being closely tied to
Hadoop which means that there was already tools available for writing
HBase-centric MapReduce jobs in Hadoop. The technology culture at Last.fm again
proved to be the main reason: there were already projects within the company
that were using HBase, so the needed infrastructure and know-how was there.

The import facility makes use of MapReduce and a custom mapper and reducer.
There is still the unresolved issue of whether it would be more efficient to not
use a reducer at all and write to HBase directly from the mapper. The argument
is that HBase is better at handling writes to random locations in the key space.
The downside of this would be that summing for aggregation would have to be done
at every cell update.

The data export uses a simple HTTP server. The query and response fit well into
the HTTP primitives and the ubiquity of HTTP clients and libraries makes it a
good choice for a simple and quick interface. The HTTP server queries HBase over
Thrift, which is quick enough but less than optimal. A more advanced solution
would be to implement a Thrift server in Java and offer a Thrift interface to
Zohmg. That would presumably be more efficient. There could also be situations
where a MapReduce job would be well-served to use the aggregates as input. That
situation would be resolved by providing an custom input format that reads
straight from the HBase table.

The decision to use JSON was based on the fact that JSON is lighweight and
efficient to parse, and has support for the few primitives that are needed to
describe the exported data. Also, JSON parsers are ubiquitious and libraries are
available in almost any language.
